Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2014-01-31 Revision/Response/response.tex

\documentclass[aoas, preprint]{imsart}

%\long\def\authornote#1{%
%        \leavevmode\unskip\raisebox{-3.5pt}{\rlap{$\scriptstyle\diamond$}}%
%        \marginpar{\raggedright\hbadness=10000
%        \def\baselinestretch{0.8}\tiny
%        \it #1\par}}
%\newcommand{\ville}[1]{\authornote{NOTE TO SELF: #1}}

%\usepackage{amsthm,amsmath,natbib}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
%\arxiv{arXiv:0000.0000}

% put your definitions there:
\startlocaldefs
\endlocaldefs

\newcommand{\argmin}{\operatornamewithlimits{arg\ min}}
\newcommand{\argmax}{\operatornamewithlimits{arg\ max}}
\newcommand\ville[1]{{\color{red} #1 }}
%    \mymarginpar{\raggedright\hbadness=10000\tiny\it #1\par}}

%\usepackage[section]{placeins}
%\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{array}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
%\usepackage{geometry}
%\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
%\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
%\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}
\usepackage[OT1]{fontenc}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\endlocaldefs

%\renewcommand{\baselinestretch}{1.2}
%\setlength{\topmargin}{-0.3in}
%\setlength{\textwidth}{6in}
%\setlength{\textheight}{8.5in}
%\setlength{\oddsidemargin}{0.25in}
%\setlength{\evensidemargin}{0.25in}
%\raggedbottom

%\doublespacing

\allowdisplaybreaks

% Math Macros.  It would be better to use the AMS LaTeX package,
% including the Bbb fonts, but I'm showing how to get by with the most
% primitive version of LaTeX.  I follow the naming convention to begin
% user-defined macro and variable names with the prefix "my" to make it
% easier to distiguish user-defined macros from LaTeX commands.
%
\newtheorem{mydef}{Definition}
\newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
\newcommand{\myZ}{\hbox{Z}^+}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\logit}{\text{logit}}
\NewDocumentCommand{\framecolorbox}{oommm}
 {% #1 = width (optional)
  % #2 = inner alignment (optional)
  % #3 = frame color
  % #4 = background color
  % #5 = text
  \IfValueTF{#1}
   {%
    \IfValueTF{#2}
     {\fcolorbox{#3}{#4}{\makebox[#1][#2]{#5}}}
     {\fcolorbox{#3}{#4}{\makebox[#1]{#5}}}%
   }
   {\fcolorbox{#3}{#4}{#5}}%
 }

\newcommand{\myfunction}[3]
{${#1} : {#2} \rightarrow {#3}$ }

\newcommand{\myzrfunction}[1]
{\myfunction{#1}{{\myZ}}{{\myR}}}

\newcommand{\mysection}[1]
{\noindent {\bf {#1}}}

%%%%%% Begin document with header and title %%%%%%%%%

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs\\
\\
(Revision of Manuscript AOAS1309-039)\\
\\
 POINT-BY-POINT RESPONSE}
\runtitle{Probability Aggregation in Time-Series}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
% \thankstext{t1}{Thanks to somebody} 
% \address{line 1\\ line 2\\ printead{e1}}
% \affiliation{Some University}

%\begin{aug}
%  \author{\fnms{First}  \snm{Author}\corref{}\thanksref{t2}\ead[label=e1]{first@somewhere.com}},
%  \author{\fnms{Second} \snm{Author}\ead[label=e2]{second@somewhere.com}}
%  \and
%  \author{\fnms{Third}  \snm{Author}%
%  \ead[label=e3]{third@somewhere.com}%
%  \ead[label=u1,url]{http://www.foo.com}}
%
%  \thankstext{t2}{Footnote to the first author with the `thankstext' command.}
%
%  \runauthor{F. Author et al.}
%
%  \affiliation{Some University and Another University}
%
%  \address{Address of the First and Second authors,\\ 
%          \printead{e1,e2}}
%
%  \address{Address of the Third author,\\
%          \printead{e3,u1}}
%
%\end{aug}

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\author{\fnms{Philip E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

\runauthor{Satop\"a\"a et al.}

 \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
 \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
 \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
\end{aug}
\end{frontmatter}

We thank both the editor and reviewers for their helpful comments on our manuscript.  We have addressed each of these comments in our revision, as well as providing a point-by-point response below. The comments are in boldface. Our comments are in normal typeface. 
\clearpage
\section*{\Large Editor}
\textbf{Concerning the out-of-sampling applicability of your methods, the cross-validation technique used in Section 6 indeed does not yield time-forward predictions, as would be required in practice.  If you could modify your technique so that it can be used for time-forward predictions, that would be most useful.  At a minimum, you should be very clear about the limitations of your approach, and you ought to discuss in detail how these issues could be remedied in future work.}\\
We have added a brief paragraph on time-forward propagation. In practice, however, social scientists tend to be more interested in the single ``best" estimate of the probability given all the information that is available at the moment. This is what our model addresses.
\\
\\
\textbf{In my own reading, I also felt that your paper could be usefully shortened.  For example, Figures 1 and 5 are redundant - so please remove Figure 1, and move Figure 5 forward to Section 1.  Also, please move the Appendix to an Online Supplement.  Throughout, please check for redundancies and remove them, unless the exposition requires otherwise.}
\\
The appendix has been moved to an Online Supplement.  Figure 1 has been replaced with Figure 5. Some sections, including the discussion on Platt calibration and summary of findings, have been removed. We also have removed many redundancies. The synthetic data section has also been shortened and now discusses only the relevant results therein.\\
\\
\textbf{As regards Referee 3's second specific comment, I agree that the definitions on page 6 are unlikely to be appreciated, unless you introduce the (rather technical) prediction space setting that is implicit in Ranjan and Gneiting (2010) and explicit in Gneiting and Ranjan ("Combining predictive distributions", Electronic Journal of Statistics 7, 1747-1782, 2013).  Therefore, let me suggest that you discuss these definitions informally (and within the text, rather than using displayed environments) and refer to the aforementioned references for technical details.}
\\
The definitions are now only discussed in the text with a reference to the aforementioned references for technical details.\\
\\
\textbf{In the references list, please unify the style used therein, capitalize all journal titles, expand acronyms such as AAAI, ECAI, and UAI, provide an update for Mellers et al. (2013), add the missing page ranges for Metropolis et al. (1953) and Niculescu-Mizil and Caruana (2005), the missing publisher for Pepe (2003), the missing authors, editors, and booktitle for Platt et al. (1999), and the multiply missing information for Wilson and Wilson (1994) and Wright et
al. (1994).}\\
- All acronyms have been expanded. We have aimed to make the style more unified.  All journal names have been capitalized.\\
- Mellers et al. (2013): This paper is currently under revision after receiving a friendly reject \& revise.\\
- Metropolis et al. (1953): The missing page ranges have been added.\\
- Niculescu-Mizil and Caruana (2005): No more referenced in the paper.\\
- Pepe (2003): Thank you for noticing this. It had been listed as an article in the .bibtex file and hence skipped the publisher info.\\
- Platt et al. (1999): No more referenced in our paper.\\
- Wilson and Wilson (1994): This has been corrected.  As this is a book, only the publisher was added.\\
- Wright et al. (1994): This reference seems to be complete.\\

\clearpage
\section*{\Large Referee 1}
\textbf{This is an interesting paper describing a model for probabilistic forecast aggregation over time.  The model is novel and has the ability to handle sparse, irregular data.  I have a few major comments and many smaller comments.\\
\\
Motivation: I think that more could be done to relate the proposed model to existing aggregation models.  As the authors state, the usual time series models do not apply to these data.  However, the proposed model can be viewed as an extension of a few psychometric models.  In particular, the model presented in Section 3 is a hierarchical factor analysis model with the traditional roles of subjects and items reversed.  Batchelder \& Romney (1988) developed a non-time-series version of this model to aggregate judgments on problems where no outcome Z exists (under the name "Cultural Consensus Theory").  Karabatsos \& Batchelder (2003) further developed Bayesian versions of this model.  Thus, if we stop after Section 3, the proposed model is an extension of cultural consensus models to time series.\\}
Thank you for pointing us to these papers. We were not aware of this work and it has definitely been an interesting read. There are so many static aggregation papers in the machine learning literature that missing some papers is unavoidable. To our understanding, the model described in Karabatsos \& Batchelder (2003) is applied to binary responses to estimate the solution key (which is also considered binary). Therefore the setup is slightly different from the probability aggregation paradigm considered in our paper. However, a more recent paper by Batchelder et al. called \textit{Cultural consensus theory: aggregating continuous responses in a finite interval} aims to use CCT to perform probability aggregation. We have cited this work in our updated introduction. This work, however, also considers static aggregation and is therefore very different in nature to our work. For this reason, and also for the sake of shortening the already too long paper, we have decided to not include an extensive literature review of the CCT. 
\\
\\
\textbf{Section 4 highlights the rotation issues inherent in factor analysis models.  The authors fit the model under one constraint, then rotate the parameters to an equivalent set of parameters that minimize a scoring rule.  So, following rotation, this model is no longer a strict consensus model because the outcomes Z impact the rotation.\\
\\
In general, I believe the relationship to factor analysis and cultural consensus provides a nice motivation for the model and its rotation issues (i.e., the calibration step).
\\
\\
Staying with the rotation issue, there are alternative constraints one may adopt instead of fixing one element of b to 1.  For example, I believe the model can be identified by restricting the sum of the elements of b to equal one (and possibly restricting the variance to 1).  This restriction may avoid the issue described on p 12, first full paragraph.}
\\
It is enough to constrain the sum to gain identifiability, i.e. no additional constraint on the variance is needed. Constraining the mean of the elements of $b$ to equal to one (or equivalently, their sum to equal to five) was considered at early stages of our implementation. 
%This implies that the groups are, on average, unbiased during the sampling step. Based on our experience this is an unreasonable assumption. 
%Unlike restricting, say, $b_3 = 1$, constraining the sum imposes unnecessary dependencies among the $b_i$'s. 
Unfortunately, the problem discussed in synthetic data section is more due to the calibration step than the choice of restriction. In essence the problem occurs because the calibration cannot move any prediction from one side of zero to the other. To see how this happens even under the sum restriction, notice that by enforcing the bias terms to sum to, say five, implies that the groups are, on average, unbiased during the sampling step. If, however, the experts are on average very under-confident, then it is likely that the resulting estimates of the hidden states are on the wrong side of zero. This is again a problem that our one-dimensional calibration cannot fix.
\\
\\
\textbf{This is a complex model, and there were a few issues that could be described in more detail.  For example, I did not see details on how the parameters $\mu_0$ and $\sigma^2_0$ were set.  This seems to be a place where one might artificially improve model performance.  }
\\
Thank you for pointing this out. In our studies we always set $\mu_0$ and $\sigma^2_0$ to 0 and 1, respectively. This is now mentioned in the text. Based on our experience, as long as these values are reasonable, they do not influence the results significantly. These values are mere baseline initializations that are rapidly washed out by the data. 
\\
\\
\textbf{Additionally, there could be further discussion about the model's handling of data sparsity.  Currently, the only details appear in the appendix (stating that an update is made for each observation at time $t$), even though data sparsity is offered as a motivation for this model.}
\\
We have added `` The first step of our Gibbs sampler is to sample the hidden states via the \textit{Forward-Filtering-Backward-Sampling} (FFBS) algorithm. FFBS first predicts the hidden states using a Kalman filter and then performs a backward sampling procedure that treats these predicted states as additional observations (see, e.g., \cite{carter1994gibbs, migon2005dynamic} for details on FFBS). Given that the Kalman filter can handle varying numbers or even no forecasts at different time points, it plays a very crucial role in our probability aggregation under sparse data."  Without describing the technical details of the Kalman Filter in the main text, it is difficult to explain how sparse data is handled. These details, however, are still kept in the Online Supplement. 
\\
\\
\textbf{Implementation: I wonder how the models were implemented, along with the speed of the model estimation.  I guess the models may take some time to run.  It would be nice if model estimation code were shared.}\\
We have added: "Our implementation of the sampling step is written in C++ and runs fast. For instance, to obtain a posterior sample of size 1,000 for 50 questions each with 100 time points and 50 experts takes about 215 seconds on the first author's computer (1.7 GHz Intel Core i5). "  The calibration step is currently run separately in R once the samples have been obtained from the sampling step. Given that the calibration requires a univariate optimization, it can be performed very fast. We plan to post our implementation upon publication.
\\
\\
\textbf{$BSAC_{log}$: I don't understand why we should expect this estimation method to be any different from the $SAC_{log}$ method, especially because flat priors are used.  In the Table 3 and Figure 3 results, it looks like there are some minor differences but not enough to be notable.  Hence, I am not convinced that the discussion of $BSAC_{log}$'s overconfidence is meaningful.}
\\
The main difference between $BSAC_{log}$ and $SAC_{log}$ is that $BSAC_{log}$ estimates the hidden states $\{X_{t,k}(1)\}$ and $\beta$ simultaneously (alternating between the two until convergence), whereas  $SAC_{log}$ first estimates $\{X_{t,k}(1)\}$ and then calibrates them by finding $\beta$. Therefore $BSAC_{log}$ is slightly more flexible than $SAC_{log}$. Unfortunately, it tends to overfit the data. This results into the overconfidence that it seen in our analyses. We have added ``Given that this model estimates $X_{t,k}(1)$ and $\beta$ simultaneously, it is a little more flexible than $\text{SAC}$."
\\
\\
\textbf{Finally, there were a variety of small comments and questions:\\
\\
Top p 6: I do not understand the phrase "the entire b vector is shared".  Does this mean that b has a single element?}
\\
This means that each of the elements of $b$ are considered the same across the 166 questions. We have modified the text to make this more clear. 
\\
\\
\textbf{Bottom p 9, expressing logistic regression as linear regression: I don't understand the normal error in this equation.  I would think that, if this has some equivalence to logistic regression, the error term should follow the logistic distribution.}
\\
 The error term was not needed. This entire paragraph has been removed for the sake of brevity. \\
\\
\textbf{p. 11, third line: I believe should read "$166 - N_{train}$".}
\\
This has been corrected. Thanks for catching it!
%Only a single dataset is generated under the synthetic data section. The algorithms are told the experts' forecasts and whether the events happened or not. After this we can estimate their accuracy to estimate the true hidden states, namely $X_{t,k}$. Therefore this evaluation is a little different from the real-world data section, where we do not have access to the true hidden states $X_{t,k}$ and hence must use the event indicators for performance evaluation.
\\
\\
\textbf{p. 11: Please define the EWMA method before describing results associated with it.}
\\
The definition of EWMA has been moved.\\
\\
\textbf{Figure 2: Would be nice to have common y-axis limits for each row of the graph.}
\\
This is a good idea. For the sake of brevity, however, we have ended up removing most of these plots. We only kept the one that tells something non-obvious about our model, and then briefly mention the other results within the text. After all this section is a mere sanity check.
\\
\\
\textbf{Appendix, p 26, top: I was initially confused because, e.g., e is missing an i subscript even though i appears in the equation.  I guess this is related to the text immediately below, stating that a new update is repeated for each observation.  But perhaps the notation could be clearer.}
\\
 The Appendix has been moved to Online Supplement, where the update step is not described in an algorithmic way and with improved notation. We believe that it is clearer now.

\clearpage
\section*{\Large Referee 2}

\textbf{This paper attempts a difficult problem of combining sparse expert probabilities which evolve over time. Three characteristics of the approach are significant }
\begin{itemize}
	\item \textbf{Experts are allowed to give multiple forecasts over time till the geo-political event materializes}
	\item \textbf{Experts are can update forecasts at irregular intervals of their liking during the same period}
	\item \textbf{Experts are grouped into multiple categories based on their level of self-acclaimed expertise}
\end{itemize}

### 2014-01-31 Revision/Paper/paper.tex
added revision

%\usepackage[section]{placeins}
%\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{array}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
%\usepackage{geometry}
%\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
%\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
%\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}
\usepackage[OT1]{fontenc}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. Their probability assessments are often evaluated and compared on \textit{calibration}, which measures how closely the frequency of event occurrence agrees with the assigned probabilities. For instance, the proportion of occurrences is 60\% for all those events to which a well-calibrated expert assigned a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts. In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.

Calibration by itself, however, is not sufficient for useful probability estimation. Consider a relatively stationary process, such as rain on different days in a given geographic region, where the observed frequency of occurrence in the last 10 years is 45\%. In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated. This assessment, however, can be made without any subject-matter expertise or special training. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. Therefore experts should aim to make probability assessments that are as far from the baseline as possible. The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy. From the decision-maker's perspective their advice is highly useful as it involves much information and very little risk. Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}). This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.

%One such observation is related to the aggregation of multiple probabilities. 
There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance. Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date. To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly. Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

@@

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to our real-world geopolitical forecasting database. The paper concludes with a discussion on future research directions and model limitations.

@@

The experts made updates on a very infrequent basis: the average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the resulting dataset is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the dataset and its collection see \citet{ungar2012good}.

@@

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
\label{Example1}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
\label{Example2}
}

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
\label{ExamplePlots}
\end{figure}

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively correlated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.

>>>>

%\usepackage[section]{placeins}
%\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{array}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
%\usepackage{geometry}
%\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
%\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
%\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}
\usepackage[OT1]{fontenc}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. Their probability assessments are often evaluated and compared on \textit{calibration}, which measures how closely the frequency of event occurrence agrees with the assigned probabilities. For instance, the proportion of occurrences is 60\% for all those events to which a well-calibrated expert assigned a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts. In particular, \citet{wright1994coherence} argue that higher self-reported expertise can be associated with better calibration.

Calibration by itself, however, is not sufficient for useful probability estimation. Consider a relatively stationary process, such as rain on different days in a given geographic region, where the observed frequency of occurrence in the last 10 years is 45\%. In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated. His assessment, however, can be made without any subject-matter expertise. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. Experts should aim to make probability assessments that are as far from the baseline as possible. The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy. Therefore useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).

%One such observation is related to the aggregation of multiple probabilities. 
There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance. Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}), political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts 
(see, e.g., \citet{Ranjan08, satopaa, batchelder2010cultural} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the experts' advice only once before the event resolves.

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date. To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly. Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.

@@

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate the crowd belief. The second evaluation section applies the model to our real-world geopolitical forecasting database. The paper concludes with a discussion on future research directions and model limitations.

@@

The average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the resulting dataset is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the dataset and its collection see \citet{ungar2012good}.

%\begin{figure}[h!]
%\vspace{-2em}
%\centering
%\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
%\vspace{-3.5em}
%
%\hspace{-0.5em}
%\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
%\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
%\label{Example1}
%} \hspace{0.5em}
%\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
%\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
%\label{Example2}
%}
%
%\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
%\label{ExamplePlots}
%\end{figure}

@@

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. }
\label{ExamplePlotsFinal}
\end{figure}

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1b} and \ref{Example2b} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been shaded according to the level of self-reported expertise and jittered slightly to make overlaps visible. The solid line gives the posterior mean of the calibrated crowd belief as estimated by our model. The surrounding dashed lines connect the point-wise 95\% posterior intervals. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probabilities for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$. Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ is assumed to be identical across all $K$ questions.  Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where (\ref{observedpr}) denotes the observed process, (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters fixed \textit{a priori} to 0 and 1, respectively. The error terms follow
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
Therefore the parameters of the model are $\boldsymbol{b}$, $\sigma^2_k$, $\gamma_k$, and $\tau^2_k$ for $k = 1, \dots, K$. Their prior distributions are chosen to be non-informative,  $p(\boldsymbol{b}, \sigma^2_k| \boldsymbol{X}_k) \propto \sigma^2_k$ and $p(\gamma_k, \tau^2_k| \boldsymbol{X}_k) \propto \tau^2_k$.
